Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling

### 2020-01-10 lloveras_2020.tex
graficos

EUV images are affected by instrumental stray light contamination, which can be corrected by deconvolution of the point spread function (PSF). While in the case of AIA images there is no reliable determination of its PSF, it has been well determined for EUVI-B by \citet{shearer_2012}. For this article, all EUVI images were stray-light corrected. For the details of impact in $\Nm$ and $\Tm$, refer to \citet{lloveras_ba2017}.

@@

\diego{ DONDE QUEDA MEJOR COMENTAR ESTA MEJORA?
Unlike in previous works, a 3D regularization scheme was applied, thus spurious artifacts (nduced by the previous longitude latitude regularization schemei) were minimized in the first two heights of the tomographic grid.}

We refer the reader to \citet{frazin_2009} for a detailed description of the technique, and to \citet{vasquez_2016} for a recent review on all published work based on it. Here we summarize the key points. The technique involves two consecutive procedures.

In a first step, the time series of EUV images is used to solve a SRT problem, for each EUV band independently. As a result, the 3D distribution of the so called {filter band emissivity} (FBE) is determined for each band separately. The FBE, {an emissivity-type quantity}, is defined as the wavelength integral of the coronal EUV spectral emissivity and the telescope's passband function of each EUV channel. Line-of-sight (LOS) integration of the FBE provides synthetic images that can be quantitatively compared to the real data in the time series. To find the FBE, the tomographic problem is posed a global optimization problem in which the quadratic norm of the difference between all pairs of synthetic and real images is minimized. Spatial regularization terms are also included in the objective function as to minimize high frequency spurious artifacts that are due to the specific characteristics of the SRT sparse projection matrix \citep{frazin_2009,frazin_2000}. {Due to optical depth issues that may affect LOSs} passing very close to the limb, and to the decay of intensity with radii, the 3D maps produced by DEMT analysis typically cover the height range 1.02 to 1.23 \rsun \ \citep{frazin_2009}.

@@

Aiming determination of the electron density and temperature along individual magnetic field lines, first both the thermodynamic results and the magnetic field obtained with the AWSoM model were interpolated into the DEMT grid. Then, the geometry of the field lines is determined by numerical integration of the first order differential equations  ${\rm d}r/B_r = r{\rm d}\theta/B_\theta = r\,\sin(\theta)\,{\rm d}\phi/B_\phi$, both inwards and outwards, from the specified 3D coordinates of a starting point. In order to evenly sample the whole volume spanned by the DEMT reconstructions, one starting point is selected at the center of each tomographic cell at 6 uniformly spaced heights, ranging from $1.025$ to $1.225\,\mrsun$, and every 2$^\circ$ in both latitude and longitude, for a total of $96,000$ starting points. Each traced field line is classified as ``open" if it intersects the source surface, set in this work at $r=6\,\mrsun$ (where lines become radial), or as ``closed" otherwise. In the case of closed magnetic field lines, loops {are} further classified as ``small" and ``large" upon their apex height being within or beyond the range of heights studied by DEMT, respectively.

@@

\diego{aca estoy haciendo la distincion entre la densidad y temp obtenida por demt y awsom}
For each open field line, and for each leg of the closed ones, an exponential fit was then applied to the electron density and a linear fit applied to the electron temperature. For DEMT results, the data points used were $\sqravgN(r)$ and $\Tm(r)$, in case of AWSoM the data points were $\Ne(r)$ and $\Te(r)$. The exponential and linear fit equations are described by

@@

%La temperatura en funcion de la altura a lo largo de lineas muestra comportamientos lineales. En el presente trabajo nos focalizamos en lineas con temperaturas crecientes y decrecientes en foma lineal. Es decir nos quedamos con aquellas piernas donde el coeficiente de pearson entre la temperatura y la altura (rho) cumple rho mayor 0.5 (con p-value 0.05) y de esta forma se descartan aquellas piernas en las que los resultados DEMT no pueden asegurar un cambio notable de temperatura entre la base y el apice (asi no uso la palabra isotermico y evado ese problema)

%The temperature as a function of height along legs shows linear behaviors. 
In the present work we focus on lines with increasing and decreasing temperatures in linear form. In other words, we selected those legs where the {Pearson correlation coefficient $\rho$ between temperature $T$ and radius $r$ meets $|\rhoTr| > 0.5$}. \temp{(with p-value less than 0.05)}. In this way, those legs in which the DEMT results cannot ensure a noticeable change in temperature between the base and the apex, are discarded.

The linear fit allows characterization of its variation with height by means of a characteristic temperature gradient $\dr \Tm\,[\MK/\mrsun]$ along each leg.
\diego{En realidad esta regresion lineal utiliza de por si los errores, pero luego tmb los uso para el test de hipotesis, quizas vale la pena ponerlo solo ahi.}

To measure how good the quality of the exponential fit to the electron density and the linear fit to the temperature data are, a chi-square merit function was used (see Section 15.2 \citet{recipes}). The errors were estimated in \citet{lloveras_2017}, $0.07$ MK for $\Tm$ and $4.0\,10^6 cm^{-3}$ for $\Ne$ and legs with p-value $>0.005$ were rejected.

@@

\begin{enumerate}

\item 
The leg must go through at least five tomographic grid cells with reconstructed data, and there must be at least one data point in each third of the range of heights spanned by the leg.

\item 
The {value of the Pearson correlation coefficient meets $|\rhoTr| > 0.5$.}

\item 
The p-value of both the exponential and linear fit must be less than 0.05.

\end{enumerate}

For the two targeted rotations, DEMT and AWSoM results were traced along the AWSoM field lines, which provides a way to characterize the global thermodynamic state of the inner solar corona and to compare them. To that end four different types of {legs} corresponding with four different types of magnetic structures were selected:
%\temp{Hacer explicacion de los tipos de lineas bien secos, luego en fig 3 hacer notar que implica cada una en cuanto a estructura magnetica}\\

\begin{itemize}
 \item Type 0 - Small (meaning with apex within the DEMT range of heights) closed field lines in the latitudes range $[-30\mdeg,+30\mdeg]$ and $\dr \Tm < 0$. %This class thoroughly populates the equator of the Streamer belt during a minimum solar rotation.

\item  Type I - Small closed field lines in all latitudes, with $\dr \Tm > 0$. %This class thoroughly populates the inner of the Streamer belt.

\item  Type II - Large closed field lines (meaning their apex surpasses the DEMT range of heights) in the latitudes range $[-90\mdeg,-30\mdeg] \cup [+30\mdeg,+90\mdeg]$ and $\dr \Tm > 0$. %This class is mostly populated by very large trans-equatorial field lines forming the envelope of the Streamer belt.

\item Type III - Open field lines (meaning with apex out the DEMT range of heights) in the latitudes range $[-90\mdeg,-60\mdeg] \cup [+60\mdeg,+90\mdeg]$ and $\dr \Tm > 0$.

\end{itemize}
\diego{Hice las definiciones con gradiente de temperatura para que quede mas claro, pero en realidad uso el pearson.} \notebyalbert{Charlemos los tres en persona de esto.}

>>>>

@@

EUV images are affected by instrumental stray light contamination, which can be corrected by deconvolution of the point spread function (PSF). While in the case of AIA images there is no reliable determination of its PSF, it has been well determined for EUVI-B by \citet{shearer_2012}. For this article, all EUVI images were stray-light corrected. For details of the impact on the tomographic results, refer to \citet{lloveras_ba2017}.

\diego{
Unlike in previous works, a 3D regularization scheme was applied, thus spurious artifacts (nduced by the previous longitude latitude regularization schemei) were minimized in the first two heights of the tomographic grid.}

We refer the reader to \citet{frazin_2009} for a detailed description of the technique, and to \citet{vasquez_2016} for a recent review on all published work based on it. Here we summarize the key points of the technique that involves two consecutive procedures.

In a first step, the time series of EUV images is used to solve a solar rotational tomography (SRT) problem, for each EUV band independently. As a result, the 3D distribution of the so called {filter band emissivity} (FBE) is determined for each band separately. The FBE, {an emissivity-type quantity}, is defined as the wavelength integral of the coronal EUV spectral emissivity and the telescope's passband function of each EUV channel. Line-of-sight (LOS) integration of the FBE provides synthetic images that can be quantitatively compared to the real data in the time series. To find the FBE, the tomographic problem is posed a global optimization problem in which the quadratic norm of the difference between all pairs of synthetic and real images is minimized. Spatial regularization terms are also included in the objective function as to minimize high frequency spurious artifacts that are due to the specific characteristics of the SRT sparse projection matrix \citep{frazin_2009,frazin_2000}. {Due to optical depth issues that may affect LOSs} passing very close to the limb, and to the decay of intensity with radii, the 3D maps produced by DEMT analysis typically cover the height range 1.02 to 1.23 \rsun \ \citep{frazin_2009}.

@@

Aiming determination of the electron density and temperature along individual magnetic field lines, first both the thermodynamic results and the magnetic field obtained with the AWSoM model were interpolated into the DEMT grid. Then, the geometry of the field lines is determined by numerical integration of the first order differential equations  ${\rm d}r/B_r = r{\rm d}\theta/B_\theta = r\,\sin(\theta)\,{\rm d}\phi/B_\phi$, both inwards and outwards, from the specified 3D coordinates of a starting point. In order to evenly sample the whole volume spanned by the DEMT reconstructions, one starting point is selected at the center of each tomographic cell at 6 uniformly spaced heights, ranging from $1.025$ to $1.225\,\mrsun$, and every 2$^\circ$ in both latitude and longitude, for a total of $96,000$ starting points. Each traced field line is classified as ``open" if it intersects the source surface, set in this work at $r=6\,\mrsun$ (where lines become radial), or as ``closed" otherwise. In the case of closed magnetic lines, they are separated in two ``legs", from the base to the apex.

%\diego{Of these legs, those whose size $L$ are smaller than the average lengths $\langle L \rangle$ are called small, and those with $L > \langle L \rangle$ ar called large.}
\diego{
In the case of closed magnetic field lines, legs {are} further classified as ``small" and ``large" upon their apex height being within or beyond the heigh average of the closed legs, respectively.
}

@@

For each open field line, and for each leg of the closed ones, an exponential fit was then applied to the electron density and a linear fit applied to the electron temperature. For DEMT results the data points used were $\sqravgN(r)$ and $\Tm(r)$, and in case of AWSoM the data points used were $\Ne(r)$ and $\Te(r)$. The exponential and linear fit equations are described by

@@

In the present work we focus on legs with increasing and decreasing temperatures in linear form. In other words, we selected those legs where the {Pearson correlation coefficient $\rho$ between temperature $T$ and radius $r$ meets $|\rhoTr| > 0.5$} {(with 95\% confidence level)}. In this way, those legs in which the DEMT results cannot ensure a noticeable change in temperature between the base and the apex, are discarded.

The linear fit allows characterization of its variation with height by means of a characteristic temperature gradient $\dr \Tm\,[\MK/\mrsun]$ along each leg. Both fits consider tomographic systematic uncertainty estimated in \citet{lloveras_2017}, i.e. $0.07$ MK for $\Tm$ and $4.0\,10^6 cm^{-3}$ for $\Ne$.

To measure how good the quality of the fit are, a chi-square merit function was used (Section 15.2 \citet{recipes}).%, selecting only legs with 95\% confidence level.

@@

\diego{
\begin{enumerate}
\item 
The leg must go through at least five tomographic grid cells with reconstructed data, and there must be at least one data point in each third of the range of heights spanned by the leg.
\item 
The {value of the Pearson correlation coefficient meets $|\rhoTr| > 0.5$.}
\item 
The coefficient level of both the exponential and linear fit must be grater than \%90.
\end{enumerate}
}

For the two targeted rotations, DEMT and AWSoM results were traced along the AWSoM field lines, which provides a way to characterize the global thermodynamic state of the inner solar corona and to compare them. To that end four different types of {legs} corresponding with four different types of magnetic structures were selected:
%\temp{Hacer explicacion de los tipos de lineas bien secos, luego en fig 3 hacer notar que implica cada una en cuanto a estructura magnetica}\\
\diego{
\begin{itemize}
\item Type 0 - Small legs (coming from closed field) in the latitudes range $[-50\mdeg,+50\mdeg]$, and $\rhoTr < -0.5$. 
 %This class thoroughly populates the equator of the Streamer belt during a minimum solar rotation.
\item Type I - Small legs (coming from closed field) in the latitudes range $[-50\mdeg,+50\mdeg]$, and $\rhoTr > 0.5$. 
%This class thoroughly populates the inner of the Streamer belt.
\item Type II - Large legs (coming from closed field) in the latitudes range $[-90\mdeg,-40\mdeg] \cup [+40\mdeg,+90\mdeg]$, and $\rhoTr > 0.5$. 
%This class is mostly populated by very large trans-equatorial field lines forming the envelope of the Streamer belt. 
\item Type III - Open field lines in the latitudes range $[-90\mdeg,-60\mdeg] \cup [+60\mdeg,+90\mdeg]$, and $\rhoTr > 0.5$.
\end{itemize}
}
